---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# God, women, and emotions: How present are referenced in the most prominent schools of philosophy?

### *Rodrigo Zarate / Applied Data Science @ Columbia University / Fall 2021*

```{r, include = FALSE} 
rm(list=ls())
memory.limit(size = 200000)
# Install required packages if needed ######################

if (!require("Rcpp")) install.packages("Rcpp")
if (!require("knitr")) install.packages("knitr")
if (!require("dplyr")) install.packages("dplyr")
if (!require("tidyr")) install.packages("tidyr")
if (!require("wordcloud")) install.packages("wordcloud")
if (!require("NLP")) install.packages("NLP")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("tm")) install.packages("tm")
if (!require("pacman")) install.packages("pacman")
if (!require("ggplot2")) install.packages("ggplot2")
pacman::p_load(pacman, party, psych, rio, tidyverse)

# LOAD AND PREPARE DATA ####################################

# Import CSV files with readr::read_csv() from tidyverse
df <- read_csv("data/philosophy_data.csv")

table.schools <- unique(df %>% 
                          select(school)%>%
                          group_by(school))
table.summary <- unique(df %>% 
                          select(school, author, title)%>%
                          group_by(school, author, title))

table.summary <- as.matrix(table.summary)
table.schools <- as.matrix(table.schools)

```

## Introduction

Have you ever wondered how present God's concept is in primary texts in the history of philosophy? In what schools of philosophy are more recurrent references to God? What about the presence of words referencing women? What are the schools of philosophy that most recurrent mention words describing women? How is the balance between men's and women's words? How is the balance of words referencing logic and emotions? A unique and fascinating data set allowed us to explore the great
texts in the history of philosophy.

The **Philosophy Data Project** (see <http://philosophydata.com/index.html>) is a data science project that explores the history of philosophy through the primary schools of thought. This fascinating data set allows us a data-driven exploration of philosophical
texts.

## Context

The dataset contains `r length(table.schools[,1])` schools of philosophy which give a broad overview of different ideas about the world, representing more or less the entire history of philosophy.

``` {r, layout="l-body-outset"}
kable(table.schools)

```

For each school of thought, there are specific philosophical texts chosen to represent the intellectual tradition.

``` {r, layout="l-body-outset"}
kable(table.summary)

```

## Data Exploration

Before starting the analysis of specific concepts present in each school of thought, I will give you a quick summary of the most prominent terms for each philosophical tradition though word clouds.

### Word clouds

A word cloud is a collection, or cluster, of words depicted in different sizes. The bigger and bolder the word appears, the more often it’s mentioned within a given text and the more important it is.

**Analytic**

``` {r, include = FALSE }

df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "analytic")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
analytic <- d
rm(m,v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Aristotle**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "aristotle")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
aristotle <- d
rm(m,v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Capitalism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "capitalism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
capitalism <- d
rm(m,v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(3,.4),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Communism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "communism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
communism <- d
rm(m,v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Continental**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "continental")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
continental <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Empiricism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "empiricism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
empiricism <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Feminism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "feminism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
feminism <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**German Idealism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "german_idealism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
german_idealism <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(3,.4),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Nietzsche**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "nietzsche")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
nietzsche <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Phenomenology**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "phenomenology")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
phenomenology <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Plato**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "plato")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
plato <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Rationalism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "rationalism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
rationalism <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

**Stoicism**

``` {r, include = FALSE }
df1 <- df %>% 
  select(school, author, title, sentence_lowered, original_publication_date)%>%
  filter(school == "stoicism")

# Create a corpus  
corpus <- VCorpus(VectorSource(df1$sentence_lowered))

corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removeWords, stopwords("english"))

corpus <- tm_map(corpus, removePunctuation)

corpus <- tm_map(corpus, stripWhitespace)


tdm <- TermDocumentMatrix(corpus)
rm(corpus, df1, d)

m <- as.matrix(tdm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(words = names(v), freq=v)
stoicism <- d
rm(m, v)
```

``` {r, echo = FALSE}
wordcloud(d$word, d$freq, random.order=FALSE, rot.per=0.3, scale=c(4,.5),
          max.words=101, colors=brewer.pal(8,"Dark2"))
```

## Logic versus Emotions

Logic and emotion are the two elements that make for perfect persuasion. We can be persuasive using only logic or emotion, but the effect will be short-term and unbalanced. In this analysis, I will show you what is the proportion of words referring to both concepts for every school of philosophy.

``` {r, include = FALSE }
women.data <- c("women","woman", "lady", "ladies",	"girl",	"girls",	"female",	"feminine")
women.data <- matrix(women.data,nrow=8)

women <- merge(women.data,analytic, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,aristotle, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,capitalism, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,communism, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,continental, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,empiricism, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,feminism, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,german_idealism, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,nietzsche, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,phenomenology, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,plato, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,rationalism, by.x="V1", by.y="words", all.x=TRUE)
women <- merge(women,stoicism, by.x="V1", by.y="words", all.x=TRUE)

names(women) <- c("word", "analytic", "aristotle", "capitalism", "communism", "continental", 
                  "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                  "plato", "rationalism", "stoicism")

men.data <- c("men","man", "gentleman", "gentlemen",	"boy",	"boys",	"male",	"mascuine")
men.data <- matrix(men.data,nrow=8)

men <- merge(men.data,analytic, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,aristotle, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,capitalism, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,communism, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,continental, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,empiricism, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,feminism, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,german_idealism, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,nietzsche, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,phenomenology, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,plato, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,rationalism, by.x="V1", by.y="words", all.x=TRUE)
men <- merge(men,stoicism, by.x="V1", by.y="words", all.x=TRUE)

names(men) <- c("word", "analytic", "aristotle", "capitalism", "communism", "continental", 
                "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                "plato", "rationalism", "stoicism")

emotions.data <- c("emotions","emotion", "feel", "love",	"passion",	"sentiment",	"sentiments",	"fear")
emotions.data <- matrix(emotions.data,nrow=8)

emotions <- merge(emotions.data,analytic, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,aristotle, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,capitalism, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,communism, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,continental, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,empiricism, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,feminism, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,german_idealism, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,nietzsche, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,phenomenology, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,plato, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,rationalism, by.x="V1", by.y="words", all.x=TRUE)
emotions <- merge(emotions,stoicism, by.x="V1", by.y="words", all.x=TRUE)

names(emotions) <- c("word", "analytic", "aristotle", "capitalism", "communism", "continental", 
                     "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                     "plato", "rationalism", "stoicism")

reasons.data <- c("reasons","logic", "think", "rational",	"coherent",	"analytic",	"analysis",	"reason")
reasons.data <- matrix(reasons.data,nrow=8)

reasons <- merge(reasons.data,analytic, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,aristotle, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,capitalism, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,communism, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,continental, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,empiricism, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,feminism, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,german_idealism, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,nietzsche, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,phenomenology, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,plato, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,rationalism, by.x="V1", by.y="words", all.x=TRUE)
reasons <- merge(reasons,stoicism, by.x="V1", by.y="words", all.x=TRUE)

names(reasons) <- c("word", "analytic", "aristotle", "capitalism", "communism", "continental", 
                    "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                    "plato", "rationalism", "stoicism")

god.data <- c("god","goddess", "divinity", "creator",	"gods",	"deity",	"deities",	"divine")
god.data <- matrix(god.data,nrow=8)

god <- merge(god.data,analytic, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,aristotle, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,capitalism, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,communism, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,continental, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,empiricism, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,feminism, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,german_idealism, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,nietzsche, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,phenomenology, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,plato, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,rationalism, by.x="V1", by.y="words", all.x=TRUE)
god <- merge(god,stoicism, by.x="V1", by.y="words", all.x=TRUE)

names(god) <- c("word", "analytic", "aristotle", "capitalism", "communism", "continental", 
                "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                "plato", "rationalism", "stoicism")


women[is.na(women)] <- 0
men[is.na(men)] <- 0
emotions[is.na(emotions)] <- 0
reasons[is.na(reasons)] <- 0
god[is.na(god)] <- 0

schools.dat <- matrix(c("analytic", "aristotle", "capitalism", "communism", "continental", 
                 "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                 "plato", "rationalism", "stoicism"), nrow=13)


women.percentage <- matrix(c(sum(women$analytic)/sum(analytic$freq), sum(women$aristotle)/sum(aristotle$freq), 
                             sum(women$capitalism)/sum(capitalism$freq),sum(women$communism)/sum(communism$freq), 
                             sum(women$continental)/ sum(continental$freq), sum(women$empiricism)/sum(empiricism$freq), 
                             sum(women$feminism)/sum(feminism$freq), sum(women$german_idealism)/sum(german_idealism$freq), 
                             sum(women$nietzsche)/sum(nietzsche$freq), sum(women$phenomenology)/sum(phenomenology$freq), 
                             sum(women$plato)/sum(plato$freq), sum(women$rationalism)/sum(rationalism$freq),
                             sum(women$stoicism)/sum(stoicism$freq)),nrow=13)

as.numeric(women.percentage)

women.percentage <- round(women.percentage*1000,2)


men.percentage <- matrix(c(sum(men$analytic)/sum(analytic$freq), sum(men$aristotle)/sum(aristotle$freq), 
                           sum(men$capitalism)/sum(capitalism$freq),sum(men$communism)/sum(communism$freq), 
                           sum(men$continental)/ sum(continental$freq), sum(men$empiricism)/sum(empiricism$freq), 
                           sum(men$feminism)/sum(feminism$freq), sum(men$german_idealism)/sum(german_idealism$freq), 
                           sum(men$nietzsche)/sum(nietzsche$freq), sum(men$phenomenology)/sum(phenomenology$freq), 
                           sum(men$plato)/sum(plato$freq), sum(men$rationalism)/sum(rationalism$freq),
                           sum(men$stoicism)/sum(stoicism$freq)),nrow=13)

as.numeric(men.percentage)

men.percentage <- round(men.percentage*1000,2)

emotions.percentage <- matrix(c(sum(emotions$analytic)/sum(analytic$freq), sum(emotions$aristotle)/sum(aristotle$freq), 
                                sum(emotions$capitalism)/sum(capitalism$freq),sum(emotions$communism)/sum(communism$freq), 
                                sum(emotions$continental)/ sum(continental$freq), sum(emotions$empiricism)/sum(empiricism$freq), 
                                sum(emotions$feminism)/sum(feminism$freq), sum(emotions$german_idealism)/sum(german_idealism$freq), 
                                sum(emotions$nietzsche)/sum(nietzsche$freq), sum(emotions$phenomenology)/sum(phenomenology$freq), 
                                sum(emotions$plato)/sum(plato$freq), sum(emotions$rationalism)/sum(rationalism$freq),
                                sum(emotions$stoicism)/sum(stoicism$freq)),nrow=13)

as.numeric(emotions.percentage)

emotions.percentage <- round(emotions.percentage*1000,2)

reasons.percentage <- matrix(c(sum(reasons$analytic)/sum(analytic$freq), sum(reasons$aristotle)/sum(aristotle$freq), 
                               sum(reasons$capitalism)/sum(capitalism$freq),sum(reasons$communism)/sum(communism$freq), 
                               sum(reasons$continental)/ sum(continental$freq), sum(reasons$empiricism)/sum(empiricism$freq), 
                               sum(reasons$feminism)/sum(feminism$freq), sum(reasons$german_idealism)/sum(german_idealism$freq), 
                               sum(reasons$nietzsche)/sum(nietzsche$freq), sum(reasons$phenomenology)/sum(phenomenology$freq), 
                               sum(reasons$plato)/sum(plato$freq), sum(reasons$rationalism)/sum(rationalism$freq),
                               sum(reasons$stoicism)/sum(stoicism$freq)),nrow=13)

as.numeric(reasons.percentage)

reasons.percentage <- round(reasons.percentage*1000,2)

god.percentage <- matrix(c(sum(god$analytic)/sum(analytic$freq), sum(god$aristotle)/sum(aristotle$freq), 
                           sum(god$capitalism)/sum(capitalism$freq),sum(god$communism)/sum(communism$freq), 
                           sum(god$continental)/ sum(continental$freq), sum(god$empiricism)/sum(empiricism$freq), 
                           sum(god$feminism)/sum(feminism$freq), sum(god$german_idealism)/sum(german_idealism$freq), 
                           sum(god$nietzsche)/sum(nietzsche$freq), sum(god$phenomenology)/sum(phenomenology$freq), 
                           sum(god$plato)/sum(plato$freq), sum(god$rationalism)/sum(rationalism$freq),
                           sum(god$stoicism)/sum(stoicism$freq)),nrow=13)

as.numeric(god.percentage)

god.percentage <- round(god.percentage*1000,2)

results <- matrix(c(schools.dat,as.numeric(women.percentage), as.numeric(men.percentage), 
                    as.numeric(emotions.percentage), as.numeric(reasons.percentage),
                    as.numeric(god.percentage)), nrow=13)

colnames(results) <- c("schools", "women", "men", "emotions", "reasons", "god")


women.table <- matrix(c(c("analytic", "aristotle", "capitalism", "communism", "continental", 
                          "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                          "plato", "rationalism", "stoicism"),
                        c("# words employed to describe WOMEN","# words employed to describe WOMEN",
                          "# words employed to describe WOMEN","# words employed to describe WOMEN",
                          "# words employed to describe WOMEN","# words employed to describe WOMEN",
                          "# words employed to describe WOMEN","# words employed to describe WOMEN",
                          "# words employed to describe WOMEN","# words employed to describe WOMEN",
                          "# words employed to describe WOMEN","# words employed to describe WOMEN",
                          "# words employed to describe WOMEN"),results[,2]), nrow=13)


colnames(women.table) <- c("School","Legend", "Value")

men.table <- matrix(c(c("analytic", "aristotle", "capitalism", "communism", "continental", 
                        "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                        "plato", "rationalism", "stoicism"),
                      c("# words employed to describe MEN","# words employed to describe MEN",
                        "# words employed to describe MEN","# words employed to describe MEN",
                        "# words employed to describe MEN","# words employed to describe MEN",
                        "# words employed to describe MEN","# words employed to describe MEN",
                        "# words employed to describe MEN","# words employed to describe MEN",
                        "# words employed to describe MEN","# words employed to describe MEN",
                        "# words employed to describe MEN"),results[,3]), nrow=13)

colnames(men.table) <- c("School","Legend", "Value")

women.men.table <- rbind(women.table,men.table)

emotions.table <- matrix(c(c("analytic", "aristotle", "capitalism", "communism", "continental", 
                             "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                             "plato", "rationalism", "stoicism"),
                           c("# words employed to describe EMOTION","# words employed to describe EMOTION",
                             "# words employed to describe EMOTION","# words employed to describe EMOTION",
                             "# words employed to describe EMOTION","# words employed to describe EMOTION",
                             "# words employed to describe EMOTION","# words employed to describe EMOTION",
                             "# words employed to describe EMOTION","# words employed to describe EMOTION",
                             "# words employed to describe EMOTION","# words employed to describe EMOTION",
                             "# words employed to describe EMOTION"),results[,4]), nrow=13)


colnames(emotions.table) <- c("School","Legend", "Value")

reasons.table <- matrix(c(c("analytic", "aristotle", "capitalism", "communism", "continental", 
                            "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                            "plato", "rationalism", "stoicism"),
                          c("# words employed to describe REASON","# words employed to describe REASON",
                            "# words employed to describe REASON","# words employed to describe REASON",
                            "# words employed to describe REASON","# words employed to describe REASON",
                            "# words employed to describe REASON","# words employed to describe REASON",
                            "# words employed to describe REASON","# words employed to describe REASON",
                            "# words employed to describe REASON","# words employed to describe REASON",
                            "# words employed to describe REASON"),results[,5]), nrow=13)

colnames(reasons.table) <- c("School","Legend", "Value")

emotions.reasons.table <- rbind(emotions.table,reasons.table)

god.table <- matrix(c(c("analytic", "aristotle", "capitalism", "communism", "continental", 
                        "empiricism", "feminism", "german_idealism", "nietzsche", "phenomenology", 
                        "plato", "rationalism", "stoicism"),
                      c("# words employed to describe GOD","# words employed to describe GOD",
                        "# words employed to describe GOD","# words employed to describe GOD",
                        "# words employed to describe GOD","# words employed to describe GOD",
                        "# words employed to describe GOD","# words employed to describe GOD",
                        "# words employed to describe GOD","# words employed to describe GOD",
                        "# words employed to describe GOD","# words employed to describe GOD",
                        "# words employed to describe GOD"),results[,6]), nrow=13)


colnames(god.table) <- c("School","Legend", "Value")

women.men.table <- as.data.frame(women.men.table)
women.table <- as.data.frame(women.table)
men.table <- as.data.frame(men.table)
emotions.table <- as.data.frame(emotions.table)
reasons.table <- as.data.frame(reasons.table)
emotions.reasons.table <- as.data.frame(emotions.reasons.table)
god.table <- as.data.frame(god.table)

women.men.table[,3] <- as.numeric(women.men.table[,3])
women.table[,3] <- as.numeric(women.table[,3])
men.table[,3] <- as.numeric(men.table[,3])
emotions.table[,3] <- as.numeric(emotions.table[,3])
reasons.table[,3] <- as.numeric(reasons.table[,3])
emotions.reasons.table[,3] <- as.numeric(emotions.reasons.table[,3])
god.table[,3] <- as.numeric(god.table[,3])
```

``` {r, echo = FALSE}
ggplot(emotions.reasons.table, aes(x = School, y = Value, fill = Legend)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = c("#3db5ff", "#104E8B")) +
  coord_flip() +
  ylim(0,15) +
  geom_text(aes(label = Value), position = position_dodge(1), 
            vjust = 0.2, hjust = -0.1, size = 3, color = "#000000") +
  labs(title = "Words employed to describe Emotion and Reason", 
       subtitle = "Proportion per 1000 words", 
       caption = "Source: Prepared by the author using data from The Philosophy Data Project",
       x = "Schools of Philosophy", y = "Proportion per 1000 words written") +
  theme(plot.title = element_text(color = "black", size = 11, face = "bold"),
        plot.subtitle = element_text(color = "black"),
        plot.caption = element_text(face = "italic", colour="black"),
        legend.title = element_text(colour = "black",  face = "bold.italic"), 
        legend.text = element_text(face = "italic", colour="black"), 
        axis.title = element_text(face = "bold", size = (11), colour = "black"),
        axis.text = element_text(colour = "black", size = (10)))
```

Surprisingly, **rationalism** is the school of thought in which the proportion of words referring to emotions and reasons is more similar (7.18 vs. 7.91 words per 1000, respectively). This similarity could be related to the fact that rationalism is a theory in which the criterion of the truth is not sensory but intellectual and deductive. Thus, its texts discuss emotions and feelings to disregard them and then base its theory on pure logic. 
On the other hand, the **analytic** school of philosophy is the thought in which the proportion of words describing emotions and reasons is more different, being the ratio of logic words 12.47 times than that of emotional expressions.

## Prominence of women

This analysis measures the proportion of words referring to women in the central philosophy books and compares this number to the ratio of terms employed to describe men.  

In the early to mid-twentieth century, women were experiencing a transformation in their role in society. With the suffragette movement and the advent of females in the workforce during the war, the role of women in society was expanding. Thus, this analysis can help us understand which schools of thought were most closely related to this phenomenon.

``` {r, echo = FALSE}
ggplot(women.table, aes(x = School, y = Value, fill = Legend)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = c("#3db5ff", "#104E8B")) +
  coord_flip() +
  ylim(0,32) +
  geom_text(aes(label = Value), position = position_dodge(1), 
            vjust = 0.2, hjust = -0.1, size = 3, color = "#000000") +
  labs(title = "Words employed to describe Women", 
       subtitle = "Proportion per 1000 words", 
       caption = "Source: Prepared by the author using data from The Philosophy Data Project",
       x = "Schools of Philosophy", y = "Proportion per 1000 words written") +
  theme(plot.title = element_text(color = "black", size = 11, face = "bold"),
        plot.subtitle = element_text(color = "black"),
        plot.caption = element_text(face = "italic", colour="black"),
        legend.title = element_text(colour = "black",  face = "bold.italic"), 
        legend.text = element_text(face = "italic", colour="black"), 
        axis.title = element_text(face = "bold", size = (11), colour = "black"),
        axis.text = element_text(colour = "black", size = (10)))
```

It is no surprise that **feminism** is the school of philosophy in which women's words are most prominent. On the other hand, **phenomenology**, the philosophical study of the structures of experience and consciousness, is the one with the lowest proportion of words describing women.

``` {r, echo = FALSE}
ggplot(women.men.table, aes(x = School, y = Value, fill = Legend)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = c("#3db5ff", "#104E8B")) +
  coord_flip() +
  ylim(0,32) +
  geom_text(aes(label = Value), position = position_dodge(1), 
            vjust = 0.2, hjust = -0.1, size = 3, color = "#000000") +
  labs(title = "Words employed to describe Women and Men", 
          subtitle = "Proportion per 1000 words", 
       caption = "Source: Prepared by the author using data from The Philosophy Data Project",
       x = "Schools of Philosophy", y = "Proportion per 1000 words written") +
  theme(plot.title = element_text(color = "black", size = 11, face = "bold"),
    plot.subtitle = element_text(color = "black"),
    plot.caption = element_text(face = "italic", colour="black"),
    legend.title = element_text(colour = "black",  face = "bold.italic"), 
    legend.text = element_text(face = "italic", colour="black"), 
    axis.title = element_text(face = "bold", size = (11), colour = "black"),
    axis.text = element_text(colour = "black", size = (10)))
```

Comparing women and men proportions, it is interesting to highlight that **feminism** is also the one with the highest proportion of men's words

## God presence

Finally, I will study references to God and divinity in the data set. This analysis is fascinating because references to God in the different schools reflect the culture and identity the beliefs of each one of the authors.

``` {r, echo = FALSE}
ggplot(god.table, aes(x = School, y = Value, fill = Legend)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = c("#3db5ff", "#104E8B")) +
  coord_flip() +
  ylim(0,15) +
  geom_text(aes(label = Value), position = position_dodge(1), 
            vjust = 0.2, hjust = -0.1, size = 3, color = "#000000") +
  labs(title = "Words employed to describe God", 
       subtitle = "Proportion per 1000 words", 
       caption = "Source: Prepared by the author using data from The Philosophy Data Project",
       x = "Schools of Philosophy", y = "Proportion per 1000 words written") +
  theme(plot.title = element_text(color = "black", size = 11, face = "bold"),
        plot.subtitle = element_text(color = "black"),
        plot.caption = element_text(face = "italic", colour="black"),
        legend.title = element_text(colour = "black",  face = "bold.italic"), 
        legend.text = element_text(face = "italic", colour="black"), 
        axis.title = element_text(face = "bold", size = (11), colour = "black"),
        axis.text = element_text(colour = "black", size = (10)))
```

**Rationalism** is the school of thought with the highest proportion of words describing God. Nevertheless, this effect is caused by the fact that some of the books spanning this philosophy are dedicated to arguing against God's existence (see <https://plato.stanford.edu/entries/spinoza/>).

Finally, **capitalism** is the school of philosophy with the lowest proportion of words referring to God (0.03 per 1000 words).
